---
output:
  pdf_document: default
  html_document: default
---



```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))

```

# Data Sets

## dnb_results

This dataset presents the results of the "diplôme national du brevet" by school, for schools in metropolitan France and for the overseas departments and regions. This data set contains 139’580 observations.

```{r}
knitr::kable(cbind(
  c(
    '**session**',
    '**school_id**',
    '**school_type**',
    '**establishment_name**',
    '**education_sector**',
    '**municipality_code**',
    '**municipality**',
    '**department_code**',
    '**department**',
    '**academy_code**',
    '**academy_name**',
    '**region_code**',
    '**region**',
    '**registered**',
    '**present**',
    '**admitted**',
    '**admitted_without**',
    '**admitted_AB**',
    '**admitted_B**',
    '**admitted_TB**',
    '**success_rate**'
  ),
  c(
    'Year of the exam session',
    'School identification number',
    'School type divided in six categories: COLLEGE, LYCEE PROFESSIONNEL, LYCEE, EREA, CFA, and AUTRE',
    'Name of the establishment',
    'Education sector categorised as public or private',
    'Municipality code',
    'Name of the municipality',
    'Department code. It is to be noted that France has 101 departements.',
    'Name of the department',
    'Academy code',
    'Name of the academy',
    'Region code. It is to be noted that France has 18 administrative, regions',
    'Name of the region',
    'Registered candidates',
    'Candidates present for the exam',
    'Candidates admitted',
    'Candidates admitted without distinction',
    'Candidates admitted with distinction "Assez Bien"',
    'Candidates admitted with distinction "Bien"',
    'Candidates admitted with distinction "Très bien"',
    '"Success rate [Present]/[Admis] as a percentage"'
  )
), col.names = c('Variable', 'Meaning')) %>% kable_styling(position = "center")
```
[Source of the data set](https://data.education.gouv.fr/explore/dataset/fr-en-dnb-par-etablissement/table/?disjunctive.session&disjunctive.numero_d_etablissement&disjunctive.denomination_principale&disjunctive.secteur_d_enseignement&disjunctive.commune_et_arrondissement&disjunctive.commune_et_arrondissement_lib_l&disjunctive.departement&disjunctive.departement_libelle&disjunctive.academie&disjunctive.academie_libelle&disjunctive.region&disjunctive.region_libelle)



## Generation 2024 data set

This data set gathers all schools which have been awarded the "Generation 2024" label.  The objective of this label, developed in view of the Paris 2024 Olympic Games, is to develop bridges between the school world and the sports movement in order to encourage young people to take part in physical activity and sport. This data set contains 6’883 observations.


```{r}
knitr::kable(cbind(
  c(
    '**region**',
    '**academy**',
    '**department**',
    '**municipality**',
    '**establishment**',
    '**school_id**',
    '**school_type**',
    '**education_sector**',
    '**postcode**',
    '**adress**',
    '**adress_2**',
    '**mail**',
    '**students**',
    '**priority_education**',
    '**city_school**',
    '**QPV**',
    '**ULIS**',
    '**SEGPA**',
    '**sport_section**',
    '**agricultural_high_school**',
    '**military_high_school**',
    '**vocational_high_school**',
    '**establishment_web**',
    '**SIREN_SIRET**',
    '**district**',
    '**ministry**',
    '**label_start_date**',
    '**label_end_date**',
    '**y_coordinate**',
    '**x_coordinate**',
    '**epsg**',
    '**precision_on_localisation**',
    '**latitude**',
    '**longitude**',
    '**position**',
    '**engaging_30_sport**'
  ),
  c(
    'Name of the region',
    'Name of the academy',
    'Name of the department',
    'Name of the municipality',
    'Name of the establishment',
    'School identification number',
    'School type',
    'Education sector categorised as public or private',
    'Postcode',
    'Address of the establishment',
    'Additional address of the establishment',
    'E-mail address of the establishment',
    'Number of students in the school',
    'Indicates whether the school is located in a priority education network (REP) or a reinforced priority education network (REP+)',
    'Indicates whether the school is part of a city school',
    'Position relative to a priority neighbourhood of the city policy. It is a policy aimed at compensating for differences in living standards with the rest of the territory.',
    '"Indicates whether the school offers a ULIS (Localized Unit for School Inclusion)"',
    '"Indicates whether the school has a SEPGA (adapted general and     vocational education sections)"',
    'Indicates whether the school has a sports section',
    'Indicates whether the school is an agricultural high school',
    'Indicates whether the school is a military high school',
    'Indicates whether the establishment is labeled "vocational high     school"',
    'Url of the description of the establishment  page on the ONISEP     website',
    '"SIREN/SIRET number of the establishment. SIREN is for Business Register Identification System in french."',
    'Name of the district to which the school is attached',
    'Ministry responsible for the institution',
    'Start date of the "generation 2024" label. Format yyyy/mm/dd',
    'End date of the "generation 2024" label. Format yyyy/mm/dd',
    'Y coordinate of the establishment, using the EPSG coordinate system',
    'X coordinate of the establishment, using the EPSG coordinate system',
    'EPSG code of the coordinate system used to locate the establishment',
    'Specification of the geographical location of the establishment',
    'Latitude',
    'Longitude',
    'Geographical position',
    'Indicates whether the institution participates in the 30 minutes of daily physical activity programme'
  )
), col.names = c('Variable', 'Meaning')) %>% kable_styling(position = "center")
```
[Source of the data set](https://data.education.gouv.fr/explore/dataset/fr-en-etablissements-labellises-generation-2024/table/?disjunctive.academie&disjunctive.departement&disjunctive.commune&sort=uai)


## Student housing Data set

This dataset records enrolment in secondary schools according to the type of accommodation for pupils: half-board, boarding school etc. This data set contains 32’096 observations.


```{r}
knitr::kable(cbind(
  c(
    '**year_back_to_school**',
    '**region**',
    '**academy**',
    '**department **',
    '**municipality**',
    '**number**',
    '**establishment_main_name**',
    '**establishment_name**',
    '**school_type**',
    '**education_sector**',
    '**students_secondary_education**',
    '**students_higher_education**',
    '**external_students_secondary_education**',
    '**half_boarders_students_secondary_education**',
    '**boarding_students_secondary_education**',
    '**external_students_higher_education**',
    '**half_board_students_higher_education**',
    '**boarding_students_higher_education**'
  ),
  c(
    'Year of the start of the school year',
    'Name of the academic region',
    'Name of the academy',
    'Name of the department',
    'Name of the municipality',
    'School identification number',
    'Main name of the establishment',
    'Name of the establishment',
    'School type ',
    'Education sector categorised as public or private',
    'Students in secondary education',
    'Number of students in higher education',
    'External students in secondary education',
    'Half-boarders in secondary education',
    'Boarding students in secondary education ',
    'External students in higher education',
    'Half-board students in higher education',
    'Boarding students in higher education'
  )
), col.names = c('Variable', 'Meaning')) %>% kable_styling(position = "center")
```
[Source of the data set](https://data.education.gouv.fr/explore/dataset/fr-en-mode-hebergement-eleves-etablissements-2d/table/?disjunctive.rentree&disjunctive.region_academique&disjunctive.academie&disjunctive.departement&disjunctive.commune&disjunctive.uai&disjunctive.type_d_etablissement&disjunctive.secteur)


## Single-parent families dataset

This data set provides information about the single-parent families in each municipality. The census is made two years in a row every five years since 2007. This data set contains 606 observations.


```{r}
knitr::kable(cbind(
  c(
    '**geocode**',
    '**libgeo**',
    '**year**',
    '**sing_par**'
  ),
  c(
    'Departmental code from INSEE',
    'Name of department',
    'Census year',
    'Number of single-parent families'


  )
), col.names = c('Variable', 'Meaning')) %>% kable_styling(position = "center")
```
[Source of the data set](https://www.observatoire-des-territoires.gouv.fr/nombre-de-familles-monoparentales)


## Covid Data set

This is a time based data set that gives us information on the COVID 19 tests and results carried out by laboratories, hospitals, pharmacists, doctors and nurses. It is updated daily. On the 30th November, the data set contained 82'394 observations.


```{r}
knitr::kable(cbind(
  c(
    '**department_code**',
    '**test_week**',
    '**educational_level**',
    '**age_group**',
    '**pop**',
    '**positive**',
    '**tested**',
    '**incidence_rate**',
    '**positivity_rate**',
    '**screening_rate**'
  ),
  c(
    'Department code',
    'Week of the tests. Format yyyy-Sww',
    'Description of the age group as [m-n], m and n being the lower and upper limits.',
    'Denomination of the age group. n-1 is used in this case excepet for the oldest group where 18 is used',
    'Population',
    'Weekly patients testing positive',
    'Weekly patients tested',
    'Incidence rate',
    'Positivity rate',
    'Screening rate'
  )
), col.names = c('Variable', 'Meaning')) %>% kable_styling(position = "center")
```
[Source of the data set](https://www.data.gouv.fr/fr/datasets/r/356bde9d-18b6-4d8b-b8bb-150cba5fba74)

**Loading of the data** 
All but the single_parent dataset are CSV files with semicolons as separators. The single_parent data set is in excel format, so we have to use read_excel. We used "skip = 4" because the document includes extra header information rows.
```{r, echo=TRUE, message=FALSE, warning = FALSE}
DNB_par_etablissement <- read_delim(here::here("data/DNB-par-etablissement.csv"), ";", escape_double = FALSE, trim_ws = TRUE)
Etablissements_labellises_generation_2024 <- read_delim(here::here("data/Etablissements-labellises-generation-2024.csv"),";", escape_double = FALSE, trim_ws = TRUE)
Hebergement_eleves_etablissements_2d <- read_delim(here::here("data/Hebergement-eleves-etablissements-2d.csv"), ";", escape_double = FALSE, trim_ws = TRUE)
insee_rp_hist_xxxx <- read_excel(here::here("data/insee_rp_hist_xxxx.xlsx"), skip = 4)
covid_sp_dep_heb_cage_scol_2022_11_30_19h01 <- read_delim(here::here("data/covid_sp_dep_heb_cage_scol_2022_11_30_19h01.csv"), ";", escape_double = FALSE, trim_ws = TRUE)
```

# Data Wrangling

We have realised that some wrangling are necessary for each data sets. We have established a checklist that we will go through for each data set. We have to :

1.  Translate the column names. As we have to rename all data sets in the same way, we have created a function. The function has a data frame and a vector as inputs. It checks if the length of the vector is correct, if so it returns a tibble with the column names renamed. Otherwise, the function gives an error message stating that the vector is not the right length. 
```{r, echo = TRUE}

rename_df <- function(df, x){
  if (ncol(df) == length(x)){
    names(df) <- c(x)
    df <- as_tibble(df)
  } else {
    stop("Vector is not the right length")
  }
}
```

2.  Make sure that all data are of the right type.
3.  Make sure that the time reference (year) are all aligned with the exam session.
4.  Add a column department_fr which will be harmonized between all data set in order to join them easily. We have also decided to keep only establishments in mainland France and Corsica. We wanted to focus on this part of France in order to get more comparable results. We note that including the results from all the overseas region would be interesting for further researches.  
5.  and more. We need to make sure that the data set does not need any further specific wrangling.


## dnb_results

1.  Translate the column names.
```{r, echo = TRUE}

dnb_colnames <- c("session", "school_id", "school_type", "establishment_name", "education_sector", "municipality_code", "municipality", "department_code", "department", "academy_code", "academy_name", "region_code", "region", "registered", "present", "admitted", "admitted_without", "admitted_AB", "admitted_B", "admitted_TB", "success_rate_pct"
)
dnb_results <- rename_df(DNB_par_etablissement, dnb_colnames)
```
2.  success_rate is of the form `xx,xx%` we want it as a double of the form `xx.xx`
```{r, echo = TRUE}

dnb_results[["success_rate_pct"]] <- as.double(gsub("%","",
                                               gsub(",",".", dnb_results[["success_rate_pct"]])))
```
3.  We need to harmonize the year variables of all the other data sets to match the logic of this one. The year is the year of the exam session (e.g academic period "2020-2021" is represented as 2021)
4.  We need to add the column `department_fr` and drop the overseas collectivities (COM).
```{r, echo = TRUE}

dnb_results$department_fr <- stri_trans_general(dnb_results$department, "Latin-ASCII") %>%
  str_to_title(.) %>% 
  gsub("Du", "du", .) %>% 
  gsub("De", "de", .) %>% 
  gsub("D'", "D", .) %>%
  gsub("Et", "et", .) %>%
  gsub(" ", "-", .) %>%
  str_replace_all("Corse-du-Sud", "Corse du Sud") %>% 
  str_replace_all("deux-Sevres", "Deux-Sevres") %>% 
  str_replace_all("Alpes-de-Hte-Provence", "Alpes-de-Haute-Provence") %>%       
  str_replace_all("Territoire-de-Belfort", "Territoire de Belfort") %>% 
  str_replace_all("Seine-Saint-denis", "Seine-Saint-Denis")

dnb_results <- dnb_results %>% 
dplyr::filter(!department_fr %in% c("Polynesie-Française","Guyane", "Martinique", "Guadeloupe", "La-Reunion", "Mayotte", "NA", "-"))

```
5.  We want to know the attribution rate of each distinction. It is needed in order to do some true comparisons between region without the same number of students.
```{r, echo = TRUE}
dnb_results <- dnb_results %>% 
  mutate(without_pct = admitted_without/admitted*100,
         AB_pct = admitted_AB/admitted*100,
         B_pct = admitted_B/admitted*100,
         TB_pct = admitted_TB/admitted*100
         )

```

6. Other data sets used for the project only go down to the department level. We therefore have to summarize all values to the department level for further merges and analyses between data sets.
```{r,echo=TRUE}
dnb_results_dep <- dnb_results %>% 
  select(session, department_code, region:TB_pct) %>% 
  group_by(department_fr, session, region) %>% 
  summarise(registered = sum(registered),
            present = sum(present),
            admitted = sum(admitted),
            admitted_without = sum(admitted_without),
            admitted_AB = sum(admitted_AB),
            admitted_B = sum(admitted_B),
            admitted_TB = sum(admitted_TB),
            without_pct = mean(without_pct, na.rm = TRUE),
            AB_pct = mean(AB_pct, na.rm = TRUE),
            B_pct = mean(B_pct, na.rm = TRUE),
            TB_pct = mean(TB_pct, na.rm = TRUE),
            success_rate_pct = mean(success_rate_pct, na.rm = TRUE))
```


We can see the summarized data set **dnb_results_dep** below. 
```{r}
datatable(dnb_results_dep, options =list(scrollX = "300px"))
```


## establishment_24

1.  Translate the column names.
```{r, echo = TRUE}
est_24_names <- c("region", "academy", "department", "municipality", "establishment", "school_id", "school_type", "education_sector", "postcode", "adress", "adress_2", "mail", "students", "priority_education", "city_school", "QPV", "ULIS", "SEGPA", "sport_section", "agricultural_high_school", "military_high_school", "vocational_high_school", "establishment_web", "SIREN_SIRET", "district", "ministry", "label_start_date", "label_end_date", "y_coordinate", "x_coordinate", "epsg", "precision_on_localisation", "latitude", "longitude", "position", "engaging_30_sport")
establishment_24 <- rename_df(Etablissements_labellises_generation_2024, est_24_names)

```
2.  No problem for this data set
3.  We need to add two variables session_started and session_ended. Indeed as the label has a start and an End date we have to trace the first session and the last session where the establishment have the label generation 2024.Most labellisations start and end in January but a few start and end in middle of the year. Exams take place end of June, beginning of July. Therefore, we will consider labellization done in August and after as done for the next academic year.
```{r, echo = TRUE}

establishment_24 <- establishment_24 %>% 
  mutate(session_started = case_when(month(label_start_date) <= 7 ~ year(label_start_date),
                                     month(label_start_date) >  7 ~ year(label_start_date)+1),
         session_ended = case_when(month(label_end_date) <= 7 ~ year(label_end_date),
                                   month(label_end_date) >  7 ~ year(label_end_date)+1)
         )
```
4.  We need to add the column `department_fr`.
```{r, echo = TRUE}

establishment_24$department_fr <- stri_trans_general(establishment_24$department, "Latin-ASCII") %>%
  str_to_title(.) %>% 
  gsub("Du", "du", .) %>% 
  gsub("De", "de", .) %>% 
  gsub("D'", "D", .) %>%
  gsub("Et", "et", .) %>%
  gsub(" ", "-", .) %>%
  str_replace_all("Corse-du-Sud", "Corse du Sud") %>% 
  str_replace_all("deux-Sevres", "Deux-Sevres") %>% 
  str_replace_all("Territoire-de-Belfort", "Territoire de Belfort") %>% 
  str_replace_all("Seine-Saint-denis", "Seine-Saint-Denis")

```

We can see on the map below, that the data set contains establishments from the overseas collectivities (COM) but from the French international schools as well.

```{r, warning=FALSE, out.width='100%', out.height= '50%'}
g <- list(
  scope = 'world',
  showland = TRUE,
  landcolor = toRGB("grey83"),
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white"),
  showlakes = TRUE,
  lakecolor = toRGB("white"),
  showsubunits = TRUE,
  showcountries = TRUE,
  resolution = 50
)

fig <- plot_geo(establishment_24, lat = ~latitude, lon = ~longitude)
fig <- fig %>% 
  layout(title = 'Establishment labellised Generation 2024', geo = g) %>% 
  add_markers(text = ~paste(department, establishment, sep = "<br />"),
              hoverinfo = "text")
fig
```
As previsouly discussed we have decided to keep only data from mainland France. We had to make sure that we also removed the French international schools. We took the opportunity to remove unused variables.
```{r, echo=TRUE}
establishment_24 <- establishment_24 %>% 
  dplyr::filter(!department_fr %in% c("Polynesie-Francaise","Guyane", "Martinique", "Guadeloupe", "La-Reunion", "Mayotte", "Saint-Martin", "-")) %>% 
  dplyr::filter(!department_fr == "NA")#"NA" and "-" makes sure that we have no more International schools. 

#establishment_24 has a lot of variables which we will for sure not use
establishment_24 <- establishment_24 %>% 
  select(-c(postcode:mail,city_school,QPV:SEGPA,establishment_web:ministry, precision_on_localisation))


```
5.  The "Diplome National du Brevet" is the diploma received at the end of "collège". We have to keep establishments which are "collège" and drop all the rest. 
```{r}
establishment_24 <- establishment_24 %>% 
  dplyr::filter(school_type == "Collège")
```
7. Other data sets only go down to the department level. We therefore have to create a simplified version for further merges between data sets. We will summarize by counting the number of establishment created each session by department. The end of the label period is not of interest for us as the first end date is on the 9.01.2021 which is categorised as session 2022 and we have results for dnb up until session 2021. 
```{r, echo=TRUE}
establishment_24_dep <- establishment_24 %>% 
  select(region, department_fr, session_started) %>% 
  group_by(department_fr, session_started) %>% 
  summarise(establishment = n())
```


We can see the summarized data set **establishment_24_dep** below. 
```{r}
datatable(establishment_24_dep, options =list(scrollX = "300px"))
```

## student_housing

1.  Translate the column names.
```{r, echo = TRUE}

housing_names <- c("year_back_to_school", "region", "academy", "department", "municipality", "school_id", "establishment_main_name", "establishment_name", "school_type", "education_sector", "students_secondary_education", "students_higher_education", "external_students_secondary_education", "half_boarders_students_secondary_education", "boarding_students_secondary_education", "external_students_higher_education", "half_board_students_higher_education", "boarding_students_higher_education")
student_housing <- rename_df(Hebergement_eleves_etablissements_2d, housing_names)

```
2.  No Problem for this data set
3.  We need to create a `session` variable as `year_back_to_school` refers to the beginning of the school year and not the exam session.
```{r, echo = TRUE}

student_housing <- student_housing %>% 
  mutate(session = year_back_to_school + 1) %>% 
    select(year_back_to_school,session, everything()) #here just to order variables
```
4.  We need to add the column `department_fr` and remove the departments outside mainland France. 
```{r, echo = TRUE}
student_housing$department_fr <- stri_trans_general(student_housing$department, "Latin-ASCII") %>%
  str_to_title(.) %>% 
  gsub("Du", "du", .) %>% 
  gsub("De", "de", .) %>% 
  gsub("D'", "D", .) %>%
  gsub("Et", "et", .) %>%
  gsub(" ", "-", .) %>%
  str_replace_all("Corse-du-Sud", "Corse du Sud") %>% 
  str_replace_all("deux-Sevres", "Deux-Sevres") %>% 
  str_replace_all("Alpes-de-Hte-Provence", "Alpes-de-Haute-Provence") %>%       
  str_replace_all("Territoire-de-Belfort", "Territoire de Belfort") %>% 
  str_replace_all("Seine-Saint-denis", "Seine-Saint-Denis")

student_housing <- student_housing %>% 
  dplyr::filter(!department_fr %in% c("Polynesie-Francaise","Guyane", "Martinique", "Guadeloupe", "La-Reunion", "Mayotte", "Saint-Martin", "-")) %>% 
  dplyr::filter(!department_fr == "NA")

```
5.  We are not interested by students from higher education and therefore can delete their variables. 
```{r, echo=TRUE}
student_housing <- student_housing %>% 
  select(-c(contains("higher")))
```
6.  We want to know the rate of each housing offering. It is needed in order to do some true comparisons between region without the same number of students.

```{r, echo=TRUE}
student_housing <- student_housing %>% 
  mutate(external_students_rate = external_students_secondary_education/students_secondary_education*100, 
         half_boarders_students_rate = half_boarders_students_secondary_education/students_secondary_education*100,
         boarding_students_rate = boarding_students_secondary_education/students_secondary_education*100)
```
7.  Other data sets only go down to the department level. We therefore have to create a simplified version for further merges between data sets. We will summarize all values to the department level.
```{r, echo=TRUE}
housing_dep <- student_housing %>% 
  select(session, region, students_secondary_education:boarding_students_rate) %>% 
  group_by(session, department_fr) %>% 
  summarise(external_students_secondary_education = sum(external_students_secondary_education, na.rm = TRUE),
            half_boarders_students_secondary_education = sum(half_boarders_students_secondary_education, na.rm = TRUE), 
            boarding_students_secondary_education = sum(boarding_students_secondary_education, na.rm = TRUE),
            students_secondary_education = sum(students_secondary_education, na.rm = TRUE),
            external_students_rate = mean(external_students_rate, na.rm = TRUE),
            half_boarders_students_rate = mean(half_boarders_students_rate, na.rm = TRUE), 
            boarding_students_rate = mean(boarding_students_rate, na.rm = TRUE))
```


We can see the summarized data set **housing_dep** below. 

```{r}
datatable(housing_dep, options =list(scrollX = "300px"))
```

## single_parent

1.  Translate the column names.
```{r , echo=TRUE}
sg_parent_names <- c("geocode", "department", "session","sing_par")
single_parent <- rename_df(insee_rp_hist_xxxx, sg_parent_names)
```
2. We have to change the scalar of the "session" and "single_par" variables into doubles.
```{r , echo=TRUE}
single_parent[["session"]]<- as.double(single_parent[["session"]])
single_parent[["sing_par"]]<- as.double(single_parent[["sing_par"]])
```
3. The session variable is already implemented in this data set.
4. We need to add the column `department_fr` and remove the department outside of mainland France. 
```{r , echo=TRUE}
single_parent$department_fr <- stri_trans_general(single_parent$department, "Latin-ASCII") %>%
  str_to_title(.) %>%
  gsub("Du", "du", .) %>%
  gsub("De", "de", .) %>%
  gsub("D'", "D", .) %>%
  gsub("Et", "et", .) %>%
  gsub(" ", "-", .) %>%
  str_replace_all("Corse-du-Sud", "Corse du Sud") %>%
  str_replace_all("deux-Sevres", "Deux-Sevres") %>%
  str_replace_all("Alpes-de-Hte-Provence", "Alpes-de-Haute-Provence") %>%      
  str_replace_all("Territoire-de-Belfort", "Territoire de Belfort") %>%
  str_replace_all("Seine-Saint-denis", "Seine-Saint-Denis")

single_parent <- single_parent %>% 
dplyr::filter(!department_fr %in% c("Polynesie-Francaise","Guyane", "Martinique", "Guadeloupe", "La-Reunion", "Mayotte", "Saint-Martin", "-")) %>% 
  dplyr::filter(!department_fr == "NA")

```
5.  No need for further data wrangling for this data set

We can see the final table **single_parent** below. 

```{r}
datatable(single_parent, options =list(scrollX = "300px"))
```

## covid_in_schools

1.  Translate the column names.
```{r, echo = TRUE}
covide_names <- c("department_code", "test_week", "educational_level", "age_group", "pop", "positive", "tested", "incidence_rate", "positivity_rate", "screening_rate")
covid_in_schools <- rename_df(covid_sp_dep_heb_cage_scol_2022_11_30_19h01,covide_names)
```
2.  test_week will be treated in (3.). positive, incidence_rate and positivity_rate need to be doubles
```{r, echo = TRUE}
covid_in_schools[["positive"]] <- as.double(gsub(",",".", covid_in_schools[["positive"]]))
covid_in_schools[["incidence_rate"]] <- as.double(gsub(",",".", covid_in_schools[["incidence_rate"]]))
covid_in_schools[["positivity_rate"]] <- as.double(gsub(",",".", covid_in_schools[["positivity_rate"]]))
```
3.  We need to create two new variables. The first will be the `test_date` categorizing each week. We chose the first day of the week. As we had the week number we had to select the week number and then for each year add seven days to the Monday of the first week of the year. The second variable is the `session`. A session is categorised from August to July of the next year. As our argument will be set on the month, we might have some test done the first days of august count towards the "wrong" session. The number of Covid cases in August are relatively low compared to the rest of the year and it represents at maximum 6 days of tests. Therefore we consider this margin of error to be satisfactory.
```{r, echo = TRUE}
covid_in_schools <- covid_in_schools %>%
  mutate(test_date = case_when (as.numeric(substr(test_week, 1,4))== 2020
                                ~ lubridate::ymd('2019-12-30') + lubridate::weeks(as.numeric(substr(test_week, 7,8))),
                                as.numeric(substr(test_week, 1,4))== 2021
                                ~ lubridate::ymd('2021-01-04') + lubridate::weeks(as.numeric(substr(test_week, 7,8))),
                                as.numeric(substr(test_week, 1,4))== 2022
                                ~ lubridate::ymd('2022-01-03') + lubridate::weeks(as.numeric(substr(test_week, 7,8)))),
         session = case_when(month(test_date) <= 7 ~ year(test_date),
                             month(test_date) >  7 ~ year(test_date)+1))
                              
```
4.  Only department code. We need to input the column `department_fr` and `region`. We use the `department_fr` and `region` variables from dnb_results. We join the two datasets by `department_code`. To do this, we first need to match the two by removing the first character of the `department_code` from dnb_results.
```{r, echo = TRUE}
reg_department <- dnb_results %>%
  select(c("department_code", "department_fr", "region")) %>%
  unique()

reg_department$department_code <- substring(reg_department$department_code, 2)
  
covid_in_schools <- right_join(x = covid_in_schools, y = reg_department, by = "department_code")
```
5.  We are only interested in the age group of 11 to 15. We use filter to keep only the students passing the DNB. 
```{r , echo=TRUE}
covid_in_schools <- covid_in_schools %>%
  filter(educational_level == "[11-15)")

```

We can see the final table *covid_in_schools* below. 
```{r, warning=FALSE}
datatable(covid_in_schools, options =list(scrollX = "300px"))
```

## Auxiliary data sets

We will use the ggplot France map for our visualizations

```{r, echo = TRUE}
map <- map_data("france")
```
The region variable is in fact the departments. We rename it "department_fr" to fit in with the other data sets.
```{r, echo = TRUE}
colnames(map)[5]<- "department_fr"

```

